---
title: "Appendix"
format:
  html:
    code-fold: true
execute: 
  echo: true
---

Import packages
```{r}
suppressWarnings(library(tidyverse))
suppressWarnings(library(cowplot))
suppressWarnings(library(knitr))
suppressWarnings(library(performance))
```

# Summary statistics
```{r}
readRDS('../data/processed/data_clean.RDS') %>% as_tibble() -> data_clean
```

# Univariate summaries of the data

## number of persons in the household
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(persons_in_household)[1], Q_1 = fivenum(persons_in_household)[2], Q_2 = fivenum(persons_in_household)[3], Q_3 = fivenum(persons_in_household)[4], Q_4 = fivenum(persons_in_household)[5]) %>% kable()
```
## number of children in the household
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(children_in_household)[1], Q_1 = fivenum(children_in_household)[2], Q_2 = fivenum(children_in_household)[3], Q_3 = fivenum(children_in_household)[4], Q_4 = fivenum(children_in_household)[5]) %>% kable()
```

## number of years of education
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(number_of_years_of_education)[1], Q_1 = fivenum(number_of_years_of_education)[2], Q_2 = fivenum(number_of_years_of_education)[3], Q_3 = fivenum(number_of_years_of_education)[4], Q_4 = fivenum(number_of_years_of_education)[5]) %>% kable()
```

## total annual income (euros)
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(total_annual_income)[1], Q_1 = fivenum(total_annual_income)[2], Q_2 = fivenum(total_annual_income)[3], Q_3 = fivenum(total_annual_income)[4], Q_4 = fivenum(total_annual_income)[5]) %>% kable()
```

## general health status
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(general_health_status)[1], Q_1 = fivenum(general_health_status)[2], Q_2 = fivenum(general_health_status)[3], Q_3 = fivenum(general_health_status)[4], Q_4 = fivenum(general_health_status)[5]) %>% kable()
```

## life satisfaction
```{r}
data_clean %>% group_by(study_year) %>% summarize(Q_0 = fivenum(life_satisfaction)[1], Q_1 = fivenum(life_satisfaction)[2], Q_2 = fivenum(life_satisfaction)[3], Q_3 = fivenum(life_satisfaction)[4], Q_4 = fivenum(life_satisfaction)[5]) %>% kable()
```

# Imputation results

* include the missingness here too

* put the imputation error estimates here. They are not the main results. 

# Model diagnostics

```{r}
readRDS('../results/lme_fit.RDS') -> lme_fit
```

```{r}
lme_fit %>% summary()
```
As can be seen in the above table, the scaled residuals are approximately normally distributed, with a slight left skew. The model `R^2` values are given below:
```{r}
performance::r2(lme_fit)
```

* put the model diagnostics *here*, including the weird fitted residuals plot, and qq plot.
* do the R^2 values go here?